### 2018-06-12 Initial Community Composition.Rmd
Update Initial Community Composition.Rmd
close, but....

---
title: "Initial Biome Composition Analysis"
author: "Timothy Brennhofer"
date: "June 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE, warning = TRUE)

#set working directory
setwd("C:/Users/Timothy Brennhofer/Desktop/KBS REU/KBS Data")
getwd()

#load necessary packages -- "library" or "require"
require(tidyverse)
require(data.table)
require(rmarkdown)
require(ggplot2)
require(dplyr)

#load data
abun_raw <- fread("2016_10_13_run2_otus.csv")
abun <- abun_raw
abun[345,4719] <- 0
abun <- abun %>%
  select(-Run, -Day, -Reads) %>%
  rename(Cage = `Cage#`) %>%
  gather(otu, abun, -SampleID, -Cage, - Abx, -Dose, -Hour, -Cohoused, -Pair) %>%
  group_by(SampleID) %>%
  mutate(abun = abun/sum(abun), otu = as.numeric(otu)) %>%
  mutate(mouseid = paste(Cage, Abx, sep = "_")) %>%
  ungroup()

```

```{r adding taxonomy}

#Load taxonomical data
tax <- fread("C:/Users/Timothy Brennhofer/Desktop/KBS REU/KBS Data/otus_table_joined_taxa.txt")

#Organize tax data into traditional classifications
tax <- tax %>%
  rename(otu = `#OTU ID`, tax = taxonomy) %>%
  mutate(tax = gsub("\\s|.__|\\[|\\]|Other", "", tax)) %>%
  separate(tax, sep = ';', c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), fill = 'right')
tax <- as.data.frame(apply(tax, 2, function(x) ifelse(x == '', 'unclassified', x)), stringsAsFactors = FALSE)
tax$otu <- as.numeric(tax$otu)

#Join taxonomical data and abundance data
abun_tax <- left_join(abun, tax, by = 'otu')

```

```{r intital composition data}

#Filter for inital reads and organize by phyla
abun_tax <- abun_tax %>%
  filter(Hour == 0) %>%
  group_by(mouseid, Phylum)

#Create graph to demonstrate initial compositions
tempz <- abun_tax %>%
  group_by(mouseid, Phylum) %>%
  group_by(mouseid) %>%
  mutate(total = sum(abun))

ggplot(tempz, aes(x = mouseid, y = total, fill = Phylum)) +
  geom_bar()

```